On The Subspace Learning for Network Attack Detection

### 2020-04-28 docs/thesis/chapters/5_conclusion_future_work/conclusion.tex
updating thesis latex

\begin{quotation}[]{Chico Xavier}
Though nobody can go back and make a new beginning, anyone can start over and make a new ending.
\end{quotation}

% falar sobre uma contextualização de forma bem resumida'

In the context of anomaly-based schemes, this thesis proposed a statistical approach based on signal processing techniques for detection of malicious traffic in computer networks. The proposed technique is based on eigenvalue analysis, model order selection (MOS) and eigen similarity analysis, where MOS and eigenvalue analysis are applied to detect time frames under attack. In addition, we evaluated the accuracy and performance of the proposed framework applied to an experimental scenario and to the DARPA 1998 data set.

This thesis also proposed a prove of concept of an architecture to evaluate the user behavior analysis on offline mobile clients through Eigensim and Model Order Selection (MOS), in order to implement an anomaly detection module based on user behavioral analysis.

Finally, we proposed the m-RPCA, which is approaches for anomaly detection on skewed and imbalanced data set, through distances of moments computed from a robust subspace learned by RPCA. The m-RPCA was evaluated for anomaly detection on simulated data and on the CTU-13 data set, which is composed by network traffic of botnets, network attacks and background flows.

\section{Conclusion}
\label{sc:conc_conclusion}

In order to be able to detect and avoid novel attacks and their variations, it is necessary to develop or improve techniques to achieve efficiency on resource consumption, processing capacity and response time. Moreover, it is crucial to obtain high detection accuracy and capacity to detect variations of malicious patterns. Recently, signal processing schemes have being applied to detect malicious traffic in computer networks \cite{Lu2009,Huang2009,Zonglin2009,david2011blind,da2012improved,tenorio2013greatest, vieira2017model}, showing advances in network traffic analysis.

This thesis models the network traffic as a signal processing formulation for applying the Eigensim, which is a framework for detection and identification of network attacks, that is based on eigenvalue analysis, model order selection (MOS) and eigen similarity analysis.

The Eigensim was evaluated and the experimental results show that synflood, fraggle and port scan attacks can be detected accurately and with great detail in an automatic and blind fashion, applying signal processing concepts for traffic modeling and through approaches based on MOS and similarity analysis of learned subspace of eigenvalues and eigenvectors. The main contributions of Eigensim were: the extension of an approach based on MOS combined with eigen analysis to blindly detect time frames under network attack; the proposal and evaluation of an eigen similarity based framework to identify details of network attacks, presenting accuracy of timely detection and identification of TCP/UDP ports under attack, as well as presenting acceptable complexity and performance regarding the processing time.

This thesis evaluated the effectiveness of MOS schemes for network attack detection, extending our previous work \cite{tenorio2013greatest} and showing that the analysis of the largest eigenvalues by time frames can be applied to detect the number of port scanning, and flood attacks, but still requiring more information for detailed attack detection. Therefore, we proposed a novel approach for detailed network attack detection, based on eigen similarity analysis.

Considering the offline mobile security context, an important issue faced by corporations that use cloud-based systems is how to provide security mechanisms to support offline corporate mobile clients. Aware of this problem and its importance, We proposed an architecture based on Eigensim for behavioral anomaly detection in offline corporate mobile apps. As prove of concept, a fully working mobile application was developed to test the proposed security solution and acquired results provide evidence that besides achieving the desired security features, the solution also has positive results in terms of performance.

Anomalies can be hard to identify and separate from normal data due to the rare occurrences of anomalies in comparison to normal events, therefore anomaly detection algorithms have to be high discriminative, robust to contamination and able to deal with the imbalanced data problem \cite{he2008learning}.

@@

\section{Contributions}
\label{sc:conc_contributions}

We analyze problems related to detection of anomalies and information security issues, and propose new approaches to improve malicious behavior detection through signal processing techniques and subspace learning. The results of the work presented in this thesis provide the following contributions:

\begin{enumerate}
	\item We proposed an approach based on eigen similarity analysis for extracting detailed information about accurate time and network ports under network attack, and evaluated the accuracy and performance of the proposed framework applied to an experimental scenario and to the DARPA 1998 dataset;
	\item We discussed the computational complexity of the proposed framework and evaluated the required processing time for tested scenarios;
	\item We proposed an architecture and techniques for offline behavioral analysis of a corporate mobile client security architecture;
	\item We discussed the processing time of the proposed framework for mobile devices;
	\item We proposed a approaches for anomaly detection on skewed and imbalanced data set, through distances of moments computed from a robust subspace learned by RPCA;
	\item We published the following papers reporting our results:
	\begin{enumerate}
		\item T. P. B. Vieira, D. F. Ten\'orio, J. P. C. L. da Costa, E. P. de Freitas, G. Del Galdo, and R. T. de Sousa J\'unior, \textit{Model Order Selection and Eigen Similarity based Framework for Detection and Identication of Network Attacks}. Journal of Networking and Computer Applications (JNCA), Vol 90, Jul 2017, Pages 26–41 \cite{vieira2017model}.
		\item T. Galibus, T. P. B. Vieira, E. P. de Freitas, R. Albuquerque, J. P. C. L. da Costa, R. T. de Sousa Jr, V. Krasnoproshin, A. Zaleski, H. Vissia, and G. del Galdo, \textit{Offline Mode for Corporate Mobile Client Security Architecture}. Mobile Networks and Applications, Springer, Mar 2017, Pages 1-17 \cite{galibus2017offline}.
		\item K. H. C. Ramos, R. T. Sousa Jr, T. P. B. Vieira, and J. P. C. L. da Costa, \textit{Discovering Critical Success Factors for Information Technologies Governance through Bibliometric Analysis of Research Publications in This Domain}. Information (Yamaguchi), 2016 \cite{ramos2016information}.
		\item K. H. C. Ramos, T. P. B. Vieira, J. P. C. L. da Costa, and R. T. Sousa Jr., \textit{Análise Multidimensional de Fatores Críticos de Sucesso em Governança de TI na Administração Pública Federal à Luz dos Dados de Controle Externo}. in Proc. 12th International Conference on Management of Technology and Information Systems (CONTECSI), 2015, São Paulo, Brazil \cite{ramos2015}.
	\end{enumerate}
\end{enumerate}

\section{Future Work}
\label{sc:conc_futurework}

Because of time constraints imposed on the Ph.D. degree, this thesis addresses some problems, but some problems are still open and others are emerging from current results. Thus, the following issues should be investigated as future work:

\begin{itemize}
	\item Improvements for obtaining better false positive rates, as well as to make the proposed framework able to identify sparse probe attacks or subtle behaviors, such as exfiltration or covert communication, considering the evaluation of a flow-based analysis and novel data sets;
	\item Distributed or parallel processing can also be evaluated to analyze the scalability and processing capacity for monitoring high throughput network traffic;
	\item Evaluate the application of the proposed approaches to different anomaly detection problems, considering cases that are aware to behavioral analysis;
	\item Enhance the m-RPCA approaches to estimate the contamination or to not rely on previously defined contamination for anomaly detection;
	\item The m-RPCA can also be extended to be online and to learn new subspace in a adaptive fashion, by means of new robust subspace learning from new observations or through robust subspace tracking.
\end{itemize}

>>>>

% falar sobre uma contextualização de forma bem resumida'
In the context of anomaly-based schemes, in thesis  we propose propose a statistical approach based on signal processing techniques for detection of malicious traffic in computer networks. The proposed technique is based on eigenvalue analysis, model order selection (MOS) and eigen similarity analysis, where MOS and eigenvalue analysis are applied to detect time frames under attack. In addition, we evaluated the accuracy and performance of the proposed framework applied to an experimental scenario and to the DARPA 1998 data set.

We propose the m-RPCA, which an approach for anomaly detection on skewed and imbalanced data set, through distances of moments computed from a robust subspace learned by RPCA. The m-RPCA was evaluated for anomaly detection on simulated data and on the CTU-13 data set, which is composed by network traffic of botnets, network attacks and background flows.

Finally, it is proposed a proof of concept of an architecture to evaluate the user behavior analysis on offline mobile clients through Eigensimilarity and Model Order Selection (MOS), in order to implement an anomaly detection module based on user behavioral analysis.

This chapter is organized as follows. In Section \ref{sec:5_conc_conclusion}, we discuss the conclusion remarks, in Section \ref{sec:5_conc_contributions} it is present the main contributions and in \ref{sec:5_conc_futurework} we propose opportunities for future work.

\section{Conclusion}
\label{sec:5_conc_conclusion}

In order to be able to detect and avoid novel attacks and their variations, it is necessary to develop or improve techniques to achieve efficiency on resource consumption, processing capacity and response time. Moreover, it is crucial to obtain high detection accuracy and capacity to detect variations of malicious patterns. Recently, signal processing schemes have being applied to detect malicious traffic in computer networks, showing advances in network traffic analysis.

This thesis models the network traffic as a signal processing formulation for applying the Eigensimilarity, which is a framework for detection and identification of network attacks, that is based on eigenvalue analysis, model order selection (MOS) and eigen similarity analysis.

The Eigensimilarity was evaluated and the experimental results show that synflood, fraggle and port scan attacks can be detected accurately and with great detail in an automatic and blind fashion, applying signal processing concepts for traffic modeling and through approaches based on MOS and similarity analysis of learned subspace of eigenvalues and eigenvectors. The main contributions of Eigensimilarity were: the extension of an approach based on MOS combined with eigen analysis to blindly detect time frames under network attack; the proposal and evaluation of an eigen similarity based framework to identify details of network attacks, presenting accuracy of timely detection and identification of TCP/UDP ports under attack, as well as presenting acceptable complexity and performance regarding the processing time.

This thesis evaluated the effectiveness of MOS schemes for network attack detection, extending our previous work and showing that the analysis of the largest eigenvalues by time frames can be applied to detect the number of port scanning, and flood attacks, but still requiring more information for detailed attack detection. Therefore, we proposed a novel approach for detailed network attack detection, based on eigen similarity analysis.

Considering the offline mobile security context, an important issue faced by corporations that use cloud-based systems is how to provide security mechanisms to support offline corporate mobile clients. Aware of this problem and its importance, we proposed an architecture based on Eigensimilarity for behavioral anomaly detection in offline corporate mobile apps. As proof of concept, a fully working mobile application was developed to test the proposed security solution and acquired results provide evidence that besides achieving the desired security features, the solution also has positive results in terms of performance.

Anomalies can be hard to identify and separate from legitimate data due to the rare occurrences of anomalies in comparison to legitimate events, therefore anomaly detection algorithms have to be high discriminative, robust to contamination and able to deal with the imbalanced data problem.

@@

\section{Contributions}
\label{sec:5_conc_contributions}

We analyze problems related to detection of anomalies and information security issues, and propose new approaches to improve malicious behavior detection through signal processing techniques and subspace learning. The results of the work presented in this thesis provided the following contributions:

\begin{enumerate}
	\item We proposed an approach based on eigen similarity analysis for extracting detailed information about accurate time and network ports under network attack, and evaluated the accuracy and performance of the proposed framework applied to an experimental scenario and to the DARPA 1998 data set;
	\item We discussed the computational complexity of the proposed framework and evaluated the required processing time for tested scenarios;
	\item We proposed an architecture and techniques for offline behavioral analysis of a corporate mobile client security architecture;
	\item We discussed the processing time of the proposed framework for mobile devices;
	\item We proposed approaches for network anomaly detection on skewed and imbalanced data set, through distances of moments computed from a robust subspace learned by RPCA;
\end{enumerate}

\section{Future Work}
\label{sec:5_conc_futurework}

This thesis addresses some problems, but some problems are still open and others are emerging from current results. Thus, the following issues should be investigated as future work:

\begin{itemize}
	\item Improvements for obtaining better false positive rates, as well as to make the Eigensimilarity able to identify sparse probe attacks or subtle behaviors, such as exfiltration or covert communication, considering the evaluation of a flow-based analysis and novel data sets;
	\item Evaluate the application of the proposed approaches to different anomaly detection problems, considering cases that are aware to behavioral analysis;
	\item Enhance the m-RPCA approaches to estimate the contamination or to not rely on previously defined contamination for anomaly detection;
	\item The m-RPCA can also be extended to be online and to learn new subspace in an adaptive fashion, by means of new robust subspace learning from new observations or through robust subspace tracking;
	\item Evaluate improvements given by a multidimensional modeling and adoption of tensor-based approaches for subspace learning, in order to analyze complex patterns that can be revealed by multidimensional analysis and tensor-based decomposition.
\end{itemize}

### 2020-04-28 docs/thesis/chapters/appendix/a_pca_svm_cfa/pca_svm_cfa.tex
updating thesis latex

The identification of the most significant information for one problem is referred to as feature selection by the signal processing and data mining areas, as well as it can be formulated as a principal component problem, which is a widely adopted signal processing technique for data visualization and feature extraction. Feature selection aims to select a subset of relevant information from a larger dataset, in order to improve: data visualization and data understanding, storage requirements, dimensionality, processing time, discriminative sensing, and to overcome overfitting problems to improve prediction and classification performance \cite{chandrashekar2014survey}.

@@

We propose a critical factors analysis based on Principal Component Analysis (PCA) for visual discriminant analysis and based on RFE combined with Support Vector Machine (SVM) \cite{hearst1998support}, applied to the survey that evaluates the IT governance of Brazilian public organizations, in order to identify the CSF for IT governance of the public sector according to TCU. Results show how PCA can make the data discriminative and that SVM is the classifier that best performs and obtains an accuracy of 91.42\% to learn and classify according to TCU's IT governance evaluation of Brazilian public sector. Finally, SVM is used to highlight the more significant features identified by RFE, which are similar to CSFs previously identified by a qualitative analysis of the same dataset.

This chapter is organized as follows. In Section \ref{sec:b_relatedworks}, related works are discussed. Section \ref{sec:b_datamodel} presents the data model and the evaluated datasets. Section \ref{sec:b_csf_fs} describes the proposed approach for critical success factors analysis. Section \ref{sec:b_experimentalresults} discusses the experimental validation and presents the results, and Section \ref{sec:b_conclusion} draws the conclusions.

@@

Ramos \emph{et al} \cite{ramos2016information} propose an overview regarding the evolution of scientific research on IT Governance critical success factors within the domain of public administration. By means of bibliometric analysis it was investigated seminal works regarding this theme, considering the characteristic key words found during our analysis. The results present 64 critical success factors with high impact on IT governance.

Guyon \emph{et al} \cite{guyon2002gene} propose a method of gene selection utilizing Support Vector Machine methods based on Recursive Feature Elimination (RFE) and demonstrate that the selected genes yield better classification performance and are biologically relevant to cancer.  The proposed method eliminates gene redundancy automatically and yields better and more compact gene subsets.

@@

The Brazilian Federal Court of Accounts (TCU, in Portuguese) surveys data regarding IT practices of Brazilian public organizations in order to audit IT governance. The dataset with a consolidate view about the answers for this survey the IT governance index is called iGovTI. The iGovTI is composed by 201 multiple choice questions, used for ranking according to their IT governance, submitted to 349 organizations. The TCU computes the IT governance index and classifies the IT governance of each organization. Additionally, Ramos \emph{et al} \cite{ramos2016information} classifies each question regarding its relevance for IT governance through a qualitative analysis, and identify the CSFs for selected IT managers regarding IT governance.

@@

PCA is a statistical technique commonly used for signal denoising, blind source separation, data compression, data visualization, feature extraction and dimensionality reduction, where a reduced number of features is extracted retaining as much information as possible \cite{jolliffe1986principal}. It uses an orthogonal transformation to convert a set of correlated variables into a set of linearly uncorrelated variables, where the first principal components have the largest variance.

@@

Considering that PCA combines attributes and creates new ones, with measurements from all of the original variables, it is hard to identify what original variables are most relevant. Therefore, it is still necessary an additional method to reveal what are the CSF for iGovTI. For this problem, we propose a CSF analysis based on RFE.

@@

Initially, it is necessary to identify one classifier that can identify the iGovTI classification of one organization according to a training dataset. Therefore, we propose an algorithm evaluation in order to identify which one presents the best accuracy for iGovTI classification. The selected algorithm should be used by RFE to identify the CSF, that should be compared to results of the CSF qualitative analysis conducted by Ramos \emph{et al} \cite{ramos2016information}.

Feature selection doesn't combine attributes, as PCA, but just evaluates their informative quality, predictive power and select the best set. Given an external estimator or classifier, that assigns weights to features, RFE is able to select features by recursively considering smaller and smaller sets of features. First, the estimator is trained on the initial set of features and the importance of each feature is obtained either through a coefficient attribute or through a feature importances attribute. Then, the least important features are pruned from current set of features. That procedure is recursively repeated on the pruned set until the desired number of features to select is eventually reached. RFE can also rank all features according to when they were eliminated.

The variables with the least effect on training errors or largest weights indicates more relevance for a classifier. Therefore, we propose to assume that the selected most relevant features are the CSF for iGovTI and also propose to validate this assumption against the results of the CSF qualitative analysis conducted by Ramos \emph{et al} \cite{ramos2016information}.

@@

For the visual discriminant analysis we initially we compute the zero mean and the the covariance matrix $\mathbf{C}$ is computed from $\mathbf{X}$.The results are shown by Figure \ref{fig:ch2_fig1}.

\begin{figure}[h!]
     \centering 
     \includegraphics[height=10cm, width=10cm]{figs/b_csf_pca_svm/raw_igovti_covariance.eps}
     \caption{Covariance matrix of iGovTI questions.}
     \label{fig:ch2_fig1}
\end{figure}

Figure \ref{fig:ch2_fig1} presents the covariance matrix of 201 questions of iGovTI, where is possible to observe low covariance for the majority questions and high variance for just a few questions.

In the next step the eigenvectors $\mathbf{v}_i$ and eigenvalues $\lambda_i$ of $\mathbf{C}$ are computed through SVD. Considering that the first principal components have the largest variance indicated by their eigenvalues, it is necessary to evaluate the variance distribution of the measured eigenvalues. Therefore, we present the Figure \ref{fig:ch2_fig2}, wich shows the variance ECDF of the iGovTI questions and shows that 97\% of the questions have variance lesser than 2, while just 3\% of the questions have variance between 2 and 11. This result indicates that just a few principal components concentrates the most significant information and motivates the evaluation regarding the relationship between the principal components and the iGovTI classification.

\begin{figure}[h!]
     \centering 
     \includegraphics[height=8cm, width=11cm]{figs/b_csf_pca_svm/raw_variance_ecdf.eps}
     \caption{Empirical CDF of variance.}
     \label{fig:ch2_fig2}
\end{figure}

Finally, we select the two largest eigenvalues and their correspondent components in order to evaluate the relationship between the iGovTI classification for 349 organizations and the two most informative variables. The Figure \ref{fig:ch2_fig3} shows the scatter diagram that plots the organizations with higher iGovTI with larger circumferences and colors near of red, while organizations with lower iGovTI have colors near of blue and shorter circumferences. It is possible to observe that organizations with higher iGovTI are concentrated in the top left quadrant, with a visual separation of organizations with lower iGovTI, but without a clear distance separating the classes.

\begin{figure}[h!]
     \centering 
     \includegraphics[height=8cm, width=11cm]{figs/b_csf_pca_svm/raw_igovti_ranking_pc2.eps}
     \caption{iGovTI ranking from 2 principal components.}
     \label{fig:ch2_fig3}
\end{figure}

@@

RFE requires a classifier or predictor to evaluate its generalization performance by recursively removing the least significant features whose deletion will have the least effect on training errors, according to the higher variance measured from the features. We perform an algorithm evaluation in order to identify the classifier with the highest accuracy for iGovTI classification. The evaluated dataset is composed of 201 questions (features) and 349 organizations (observations), where each organization is classified according to iGovTI classes, which are initial, intermediate and enhanced. We divide the dataset into train and test, with the training data equivalent to 90\% of the whole dataset, while the test has 10 \% of the whole dataset.

The Table \ref{tab:ch2_tab1} present the selected classification algorithms and the measured accuracy for iGovTI classification. According to results, SVM \cite{hearst1998support} is the algorithm with highest classification accuracy, with 91.42\%, while the second is Elastic Net \cite{zou2005regularization}, with 83.15\% of accuracy. The results for iGovTI classification corroborates previous evaluations that highlight the advantages of SVM for classifications from small datasets \cite{guyon2002gene}. Additionally, our results indicate that SVM is the best classifier for RFE, which corroborates the previous research of Guyon \emph{et al} \cite{guyon2002gene}, where is proposed SVM-RFE, that is a method of gene selection utilizing SVM methods based on RFE.

\begin{table*}[!t]
	\caption{Evaluation of classification accuracy for iGovTI}
  	\label{tab:ch2_tab1}
	\centering
	\begin{tabular}{|l|r|}
		\hline \rowcolor{Gray} Algorithm	& Mean Accuracy\\\hline
		Linear Regression \cite{draper2014applied}	&0.3608\\ \hline
		LDA \cite{martinez2001pca}	&0.6285\\ \hline
		K-Nearest Neighbors \cite{fukunaga1975branch}	&0.7142\\ \hline
		Linear SVM \cite{fan2008liblinear}	&0.7142\\ \hline
		Logistic Regression \cite{hosmer2013applied}	&0.7714\\ \hline
		SVM Regression \cite{smola2004tutorial}	&0.8268\\ \hline
		Lasso \cite{tibshirani1996regression}	&0.8274\\ \hline
		Elastic Net \cite{zou2005regularization}	&0.8315\\ \hline
		SVM \cite{hearst1998support}	&0.9142\\ \hline
	\end{tabular}
\end{table*}

The next step is to use the selected algorithm as validator for recursive feature elimination in order to identify the CSF for iGovTI. Additionally, the RFE algorithm requires the number of desired most significants features. We adopt 54 for this variable, considering that is the number of CSF identified by Ramos \emph{et al} \cite{ramos2016information} in a qualitative analysis, that is the proposed target to validate the results of the CSF identification for iGovTI.

The RFE algorithm presents an ranking according to the significance of each feature, with an accuracy of 69,9\% for CSF classification when compared to the CSF identified by Ramos \emph{et al} \cite{ramos2016information} in a qualitative analysis.

@@

Finally, SVM is used to select the more significant features identified by RFE, which are 69,9\% similar to CSFs previously identified by a qualitative analysis of the same dataset.

>>>>

The identification of the most significant information for one problem is referred to as feature selection by the signal processing and data mining areas, as well as it can be formulated as a principal component problem, which is a widely adopted signal processing technique for data visualization and feature extraction. Feature selection aims to select a subset of relevant information from a larger data set, in order to improve: data visualization and data understanding, storage requirements, dimensionality, processing time, discriminative sensing, and to overcome overfitting problems to improve prediction and classification performance \cite{chandrashekar2014survey}.

@@

We propose a critical factors analysis based on Principal Component Analysis (PCA) for visual analysis, and based on RFE combined with Support Vector Machine (SVM) \cite{hearst1998support} for classification and identification of critical success factors, applied to the survey that evaluates the IT governance of Brazilian public organizations, in order to identify the CSF for IT governance of the public sector according to TCU. Results show how PCA can help the data visualization and the discriminative visual analysis, and that SVM is the classifier that best performs on classification and obtains an accuracy of 91.42\% to learn and classify according to TCU's IT governance evaluation of Brazilian public sector. Finally, SVM is used to highlight the more significant features identified by RFE, which are similar to CSFs previously identified by a qualitative analysis of the same data set.

This chapter is organized as follows. In Section \ref{sec:b_relatedworks}, related works are discussed. Section \ref{sec:b_datamodel} presents the data model and the evaluated data sets. Section \ref{sec:b_csf_fs} describes the proposed approach for critical success factors analysis. Section \ref{sec:b_experimentalresults} discusses the experimental validation and presents the results, and Section \ref{sec:b_conclusion} draws the conclusions.

@@

Ramos \emph{et al.} \citeme{ramos2016information, ramos2015multidimensional} propose an overview regarding the evolution of scientific research on IT Governance critical success factors within the domain of public administration. By means of bibliometric analysis it was investigated seminal works regarding this theme, considering the characteristic key words found during our analysis. The results present 64 critical success factors with high impact on IT governance.

Guyon \emph{et al.} \cite{guyon2002gene} propose a method of gene selection utilizing Support Vector Machine methods based on Recursive Feature Elimination (RFE) and demonstrate that the selected genes yield better classification performance and are biologically relevant to cancer.  The proposed method eliminates gene redundancy automatically and yields better and more compact gene subsets.

@@

The Brazilian Federal Court of Accounts (TCU, in Portuguese) surveys data regarding IT practices of Brazilian public organizations in order to audit IT governance. The data set with a consolidate view about the answers for this survey the IT governance index is called iGovTI. The iGovTI is composed by 201 multiple choice questions, used for ranking according to their IT governance, submitted to 349 organizations. The TCU computes the IT governance index and classifies the IT governance of each organization. Additionally, Ramos \emph{et al.} \citeme{ramos2016information} classifies each question regarding its relevance for IT governance through a qualitative analysis, and identify the CSFs for selected IT managers regarding IT governance.

@@

PCA is a statistical technique commonly used for signal denoising, data compression, data visualization, feature extraction and dimensionality reduction, where a reduced number of features is extracted retaining as much information as possible \cite{jolliffe1986principal}. It uses an orthogonal transformation to convert a set of correlated variables into a set of linearly uncorrelated variables, where the first principal components have the largest variance.

@@

Considering that PCA combines attributes and creates new ones, with measurements from all of the original variables, it is hard to identify what original variables are most relevant. Therefore, we do not adopt PCA for classification and identification of CSF, thus it is still necessary an additional method to reveal what are the CSF for iGovTI. For this problem, we propose a CSF analysis based on RFE.

@@

Initially, it is necessary to identify one classifier that can identify the iGovTI classification of one organization according to a training data set. Therefore, we propose an algorithm evaluation in order to identify which one presents the best accuracy for iGovTI classification. The selected algorithm should be used by RFE to identify the CSF, that should be compared to results of the CSF qualitative analysis conducted by Ramos \emph{et al.} \citeme{ramos2016information}.

Feature selection doesn't combine attributes, as PCA, but just evaluates their informative quality, predictive power and select the best set. Given an external estimator or classifier, that assigns weights to features, RFE is able to select features by recursively considering smaller and smaller sets of features. First, the estimator is trained on the initial set of features and the importance of each feature is obtained either through a coefficient attribute or through a feature importance attribute. Then, the least important features are pruned from current set of features. That procedure is recursively repeated on the pruned set until the desired number of features to select is eventually reached. RFE can also rank all features according to when they were eliminated.

The variables with the least effect on training errors or largest weights indicates more relevance for a classifier. Therefore, we propose to assume that the selected most relevant features are the CSF for iGovTI and also propose to validate this assumption against the results of the CSF qualitative analysis conducted by Ramos \emph{et al.} \citeme{ramos2016information}.

@@

For the visual discriminant analysis we initially we compute the zero mean and the the covariance matrix $\mathbf{C}$ is computed from $\mathbf{X}$.The results are shown by Figure \ref{fig:b_fig1}.

\begin{figure}[h!]
     \centering 
     \includegraphics[height=10cm, width=10cm]{figures/apx_b/raw_igovti_covariance-eps-converted-to.pdf}
     \caption{Covariance matrix of iGovTI questions.}
     \label{fig:b_fig1}
\end{figure}

Figure \ref{fig:b_fig1} presents the covariance matrix of 201 questions of iGovTI, where is possible to observe low covariance for the majority questions and high variance for just a few questions.

In the next step the eigenvectors $\mathbf{v}_i$ and eigenvalues $\lambda_i$ of $\mathbf{C}$ are computed through SVD. Considering that the first principal components have the largest variance indicated by their eigenvalues, it is necessary to evaluate the variance distribution of the measured eigenvalues. Therefore, we present the Figure \ref{fig:b_fig2}, which shows the variance ECDF of the iGovTI questions and shows that 97\% of the questions have variance lesser than 2, while just 3\% of the questions have variance between 2 and 11. This result indicates that just a few principal components concentrates the most significant information and motivates the evaluation regarding the relationship between the principal components and the iGovTI classification.

\begin{figure}[h!]
     \centering 
     \includegraphics[height=8cm, width=11cm]{figures/apx_b/raw_variance_ecdf.eps}
     \caption{Empirical CDF of variance.}
     \label{fig:b_fig2}
\end{figure}

Finally, we select the two largest eigenvalues and their correspondent components in order to evaluate the relationship between the iGovTI classification for 349 organizations and the two most informative variables. The Figure \ref{fig:b_fig3} shows the scatter diagram that plots the organizations with higher iGovTI with larger circumferences and colors near of red, while organizations with lower iGovTI have colors near of blue and shorter circumferences. It is possible to observe that organizations with higher iGovTI are concentrated in the top left quadrant, with a visual separation of organizations with lower iGovTI, but without a clear distance separating the classes.

\begin{figure}[h!]
     \centering 
     \includegraphics[height=8cm, width=11cm]{figures/apx_b/raw_igovti_ranking_pc2.eps}
     \caption{iGovTI ranking from 2 principal components.}
     \label{fig:b_fig3}
\end{figure}

@@

RFE requires a classifier or predictor to evaluate its generalization performance by recursively removing the least significant features whose deletion will have the least effect on training errors, according to the higher variance measured from the features. We perform an algorithm evaluation in order to identify the classifier with the highest accuracy for iGovTI classification. The evaluated data set is composed of 201 questions (features) and 349 organizations (observations), where each organization is classified according to iGovTI classes, which are initial, intermediate and enhanced. We divide the data set into train and test, with the training data equivalent to 90\% of the whole data set, while the test has 10 \% of the whole data set.

The Table \ref{tab:b_tab1} present the selected classification algorithms and the measured accuracy for iGovTI classification. According to results, SVM \cite{hearst1998support, reis2016esprit} is the algorithm with highest classification accuracy, with 91.42\%, while the second is Elastic Net \cite{zou2005regularization}, with 83.15\% of accuracy. The results for iGovTI classification corroborates previous evaluations that highlight the advantages of SVM for classifications from small data sets \cite{guyon2002gene}. Additionally, our results indicate that SVM is the best classifier for RFE, which corroborates the previous research of Guyon \emph{et al.} \cite{guyon2002gene}, where is proposed SVM-RFE, that is a method of gene selection utilizing SVM methods based on RFE.

\begin{table*}[!t]
	\caption{Evaluation of classification accuracy for iGovTI}
  	\label{tab:b_tab1}
	\centering
	\begin{tabular}{|l|r|}
		\hline \rowcolor{Gray} Algorithm	& Mean Accuracy\\\hline
		Linear Regression \cite{draper2014applied}	&0.3608\\ \hline
		LDA \cite{martinez2001pca}	&0.6285\\ \hline
		K-Nearest Neighbors \cite{fukunaga1975branch}	&0.7142\\ \hline
		Linear SVM \cite{fan2008liblinear}	&0.7142\\ \hline
		Logistic Regression \cite{hosmer2013applied}	&0.7714\\ \hline
		SVM Regression \cite{smola2004tutorial}	&0.8268\\ \hline
		Lasso \cite{tibshirani1996regression}	&0.8274\\ \hline
		Elastic Net \cite{zou2005regularization}	&0.8315\\ \hline
		SVM \cite{hearst1998support}	&0.9142\\ \hline
	\end{tabular}
\end{table*}

The next step is to use the selected algorithm as validator for recursive feature elimination in order to identify the CSF for iGovTI. Additionally, the RFE algorithm requires the number of desired most significant features. We adopt 54 for this variable, considering that is the number of CSF identified by Ramos \emph{et al.} \citeme{ramos2016information} in a qualitative analysis, that is the proposed target to validate the results of the CSF identification for iGovTI.

The RFE algorithm presents an ranking according to the significance of each feature, with an accuracy of 69,9\% for CSF classification when compared to the CSF identified by Ramos \emph{et al.} \citeme{ramos2016information} in a qualitative analysis.

@@

Finally, SVM is used to select the more significant features identified by RFE, which are 69,9\% similar to CSFs previously identified by a qualitative analysis of the same data set.
